---
title: "Lab 3"
author: "Allison Louie"
format:
  html:
    embed-resources: true
---

```{r}
library(data.table)
library(tidyverse)
library(R.utils)
library(leaflet)
library(leaflet.providers)
library(lubridate)
```

#1 Read in the data

```{r}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = ("met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread("met_all.gz")
```

#2 Check the dimensions, headers, footers. How many columns, rows are there?

```{r}
dim(met)
```

```{r}
head(met)
```

```{r}
tail(met)
```

#3 Take a look at the variables.
```{r}
str(met)
```

#4 Take a closer look at the key variables.

```{r}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```
```{r}
summary(met$temp)
```

```{r}
summary(met$elev)
```

```{r}
summary(met$wind.sp)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

After replacing the value of 9999.0 with NA, the weather station with highest elevation is 4113 meters.

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

#5 Check the data against an external data source

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

Now we have removed any temperatures lower than the value of -15˚C. Now the minimum temperature is -3˚C.

#6 Calculate summary statistics

```{r}
elev <- met[elev==max(elev), ]
summary(elev)
```

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```

```{r}
cor(elev$temp, elev$hour, use="complete")
```

```{r}
cor(elev$wind.sp, elev$day, use="complete")
```

```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```

```{r}
cor(elev$temp, elev$day, use="complete")
```

#7 Exploratory graphs

```{r}
hist(met$elev, breaks=100)
```

```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

